### 2015-11-07 Assignment IV/Assignment4-Ditz_Schroeder-WS15-16.tex
neueste version der tex datei beschreibung der methoden hinzugefügt

\section*{Practical Assignment - \textsl{Comparing multiple alignment}}
Multiple sequence alignments (MSA) are very important to detect similarities between several 
sequences and e.g. determine whether a sequence belongs to a certain family or not. Hence, many 
different programs, such as ClustalW, were developed to calculate a MSA. We calculated a multiple 
sequence alignment for three files (BB11007, BB20002 and Prolyl-tRNA), each file contained several 
sequences, with different programs 
to compare the results. The used programs are MAFFT\cite{Katoh1}, MUSCLE\cite{Edgar1, Edgar2} and 
Clustal$\Omega$\cite{Sievers539}. The resulting MSAs were compared with a gold standard provided by 
BaliBase\cite{Bahr1}. This database provide a large number of reference MSAs as well as a program 
\grqq BaliScore\grqq , which can be used to calculate different scores to evaluate your MSA. For 
this assignment we are interested in the total column score, which is the percentage of accurately 
reconstructed columns of the reference MSA, and the sum-of-pair score, which is for BaliScore the 
percentage of pairs of aligned residues that are similar in the reference and the reconstructed 
MSA. Furthermore, a important information is the runtime of different programs. We used the bash 
command \textit{time} to collect this information.\\
\\
\noindent The resulting total column scores can be found in figure \ref{totCol}. All calculated sum-
of-pairs scores are written down in figure \ref{SPscore} and the runtime of all programs can be found 
in figure \ref{runtime}.

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
  \textit{total column score} & \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & $c_{1,1}$ & $c_{2,1}$ & $c_{3,1}$ \\
  \textbf{BB20002} & $c_{1,2}$ & $c_{2,2}$ & $c_{3,2}$ \\
  \textbf{Prolyl-tRNA} & $c_{1,3}$ & $c_{2,3}$ & $c_{3,3}$ \\
 \end{tabular}}
 \caption{total column score of all used programs for example files}
 \label{totCol}
\end{figure}

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
   \textit{SP-score}& \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & $s_{1,1}$ & $s_{2,1}$ & $s_{3,1}$ \\
  \textbf{BB20002} & $s_{1,2}$ & $s_{2,2}$ & $s_{3,2}$ \\
  \textbf{Prolyl-tRNA} & $s_{1,3}$ & $s_{2,3}$ & $s_{3,3}$ \\
 \end{tabular}}
 \caption{sum-of-pairs score of all used programs for example files}
 \label{SPscore}
\end{figure}

>>>>

\section*{Practical Assignment - \textsl{Comparing multiple alignment}}
Multiple sequence alignments (MSA) are very important to detect similarities between several 
sequences and e.g. determine whether a sequence belongs to a certain family or not. Hence, many 
different programs, such as ClustalW, were developed to calculate a MSA. We calculated a multiple 
sequence alignment for three files (BB11007, BB20002 and Prolyl-tRNA), each file contained several 
sequences, with different programs 
to compare the results. The used programs are Clustal$\Omega$\cite{Sievers539}, MAFFT\cite{Katoh1} 
and MUSCLE\cite{Edgar1, Edgar2}. The resulting MSAs were compared with a gold standard provided by 
BaliBase\cite{Bahr1}. This database provide a large number of reference MSAs as well as a program 
\grqq BaliScore\grqq , which can be used to calculate different scores to evaluate your MSA. For 
this assignment we are interested in the total column score, which is the percentage of accurately 
reconstructed columns of the reference MSA, and the sum-of-pair score, which is for BaliScore the 
percentage of pairs of aligned residues that are similar in the reference and the reconstructed 
MSA. Furthermore, a important information is the runtime of different programs. We used the bash 
command \textit{time} to collect this information.\\
\\
\noindent Clustal$\Omega$ is the newest extension for the Clustal family. One problem of ClustalW 
was the scalability. With Clustal$\Omega$ a MSA can now calculated for hundreds of thousands of 
sequences within a few hours. On the technical side the algorithm can use multiple processors. 
According to the official site\footnote{http://www.clustal.org/omega/}, Clustal$\Omega$ produces 
better alignments than previous versions. MAFFT offers several modes. The L-INS-i mode is 
slower and only suitable for alignments with less than 200 sequences but results in a more accurate 
alignment. Contrary to that mode the FFT-NS-2 mode can be used to align up to 300.000 sequences. It 
is a lot faster by using the Fast Fourier Transformation but also less accurate. We used the 
FFT-NS-2 mode for this assignment. The third program MUSCLE is a iterative approach to multiple 
sequence alignments. It uses a very accurate distance measure two calculate relatedness of two 
sequences and updates this distance measure for every iteration step.\\
\\
\noindent The resulting total column scores can be found in figure \ref{totCol}. All calculated sum-
of-pairs scores are written down in figure \ref{SPscore} and the runtime of all programs can be found 
in figure \ref{runtime}.

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
  \textit{total column score} & \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & 0.33 & 0.39 & 0.24 \\
  \textbf{BB20002} & 0.00 & 0.00 & 0.00 \\
  \textbf{Prolyl-tRNA} & $c_{1,3}$ & $c_{2,3}$ & $c_{3,3}$ \\
 \end{tabular}}
 \caption{total column score of all used programs for example files}
 \label{totCol}
\end{figure}

\begin{figure}[ht]
 \centering
 \subfigure{
 \begin{tabular}{c|ccc}
   \textit{SP-score}& \textbf{Clustal$\Omega$} & \textbf{MUSCLE} & \textbf{MAFFT} \\
  \hline
  \textbf{BB11007} & 0.597 & 0.582 & 0.524 \\
  \textbf{BB20002} & 0.314 & 0.197 & 0.246 \\
  \textbf{Prolyl-tRNA} & $s_{1,3}$ & $s_{2,3}$ & $s_{3,3}$ \\
 \end{tabular}}
 \caption{sum-of-pairs score of all used programs for example files}
 \label{SPscore}
\end{figure}
